---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(glue)
```

# glue
[![Travis-CI Build Status](https://travis-ci.org/tidyverse/glue.svg?branch=master)](https://travis-ci.org/tidyverse/glue)
[![Coverage Status](https://img.shields.io/codecov/c/github/tidyverse/glue/master.svg)](https://codecov.io/github/tidyverse/glue?branch=master)

Glue strings to data in R. Small, fast, dependency free interpreted string literals.

## Installation

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("tidyverse/glue")
```

## Usage

Long strings can be broken by line and will be concatenated together
```{r}
name <- "Fred"
age <- 50
anniversary <- as.Date("1991-10-12")
to('My name is {name},',
  ' my age next year is {age + 1},',
  ' my anniversary is {format(anniversary, "%A, %B %d, %Y")}.')
```

You can use named arguments to assign temporary variables

```{r}
to('My name is {name},',
  ' my age next year is {age + 1},',
  ' my anniversary is {format(anniversary, "%A, %B %d, %Y")}.',
  name = "Joe",
  age = 40,
  anniversary = as.Date("2001-10-12"))
```

`to_data()` is useful in magrittr pipes
```{r}
library(magrittr)
mtcars %>% to_data("{rownames(.)} has {hp} hp")
```

Leading whitespace and blank lines are automatically trimmed, which lets you indent the strings naturally.
```{r}
fun <- function() {
  cat(to("
    A Formatted string
    Can have multiple lines
      with additional indention preserved
    "))
}

fun()
```

A literal brace can be inserted by using doubled braces.
```{r}
name <- "Fred"
to("My name is {name}, not {{name}}.")
```

All valid R code works in expressions, including braces and escaping.
Backslashes do need to be doubled just like in all R strings.
```{r}
  `foo}\`` <- "foo"
to("{
      {
        '}\\'' # { and } in comments, single quotes
        \"}\\\"\" # or double quotes are ignored
        `foo}\\`` # as are { in backticks
      }
  }")
```

# Other implementations

Some other implementations of string interpolation in R (although not using identical syntax).

- [stringr::str_interp](http://stringr.tidyverse.org/reference/str_interp.html)
- [pystr::pystr_format](https://cran.r-project.org/package=pystr)
- [R.utils::gstring](https://cran.r-project.org/package=R.utils)
