---
output:
  github_document:
    html_preview: false
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(glue)
```

# glue
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/glue)](http://cran.r-project.org/package=glue)
[![Travis-CI Build Status](http://travis-ci.org/tidyverse/glue.svg?branch=master)](http://travis-ci.org/tidyverse/glue)
[![Coverage Status](http://img.shields.io/codecov/c/github/tidyverse/glue/master.svg)](http://codecov.io/github/tidyverse/glue?branch=master)
[![AppVeyor Build Status](http://ci.appveyor.com/api/projects/status/github/tidyverse/glue?branch=master&svg=true)](http://ci.appveyor.com/project/tidyverse/glue)

Glue strings to data in R. Small, fast, dependency free interpreted string literals.

## Installation

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("tidyverse/glue")
```

## Usage

##### Long strings are broken by line and concatenated together.
```{r}
library(glue)

name <- "Fred"
age <- 50
anniversary <- as.Date("1991-10-12")
glue('My name is {name},',
  ' my age next year is {age + 1},',
  ' my anniversary is {format(anniversary, "%A, %B %d, %Y")}.')
```

##### Named arguments are used to assign temporary variables.

```{r}
glue('My name is {name},',
  ' my age next year is {age + 1},',
  ' my anniversary is {format(anniversary, "%A, %B %d, %Y")}.',
  name = "Joe",
  age = 40,
  anniversary = as.Date("2001-10-12"))
```

##### `glue_data()` is useful with [magrittr](http://cran.r-project.org/package=magrittr) pipes.
```{r}
`%>%` <- magrittr::`%>%`
head(mtcars) %>% glue_data("{rownames(.)} has {hp} hp")
```

##### Leading whitespace and blank lines from the first and last lines are automatically trimmed.
This lets you indent the strings naturally in code.
```{r}
glue("
    A formatted string
    Can have multiple lines
      with additional indention preserved
    ")
```

##### An additional newline can be used if you want a leading or trailing newline.
```{r}
glue("

  leading or trailing newlines can be added explicitly

  ")
```

##### `\\` at the end of a line continues it without a new line.
```{r}
glue("
    A formatted string \\
    can also be on a \\
    single line
    ")
```

##### A literal brace is inserted by using doubled braces.
```{r}
name <- "Fred"
glue("My name is {name}, not {{name}}.")
```

##### Alternative delimiters can be specified with `.open` and `.close`.
```{r}
one <- "1"
glue("The value of $e^{2\\pi i}$ is $<<one>>$.", .open = "<<", .close = ">>")
```

##### All valid R code works in expressions, including braces and escaping.
Backslashes do need to be doubled just like in all R strings.
```{r}
  `foo}\`` <- "foo"
glue("{
      {
        '}\\'' # { and } in comments, single quotes
        \"}\\\"\" # or double quotes are ignored
        `foo}\\`` # as are { in backticks
      }
  }")
```

##### `glue_sql()` makes constructing SQL statements safe and easy
Use backticks to quote identifiers, normal strings and numbers are quoted
appropriately for your backend.

```{r}
library(glue)

con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
colnames(iris) <- gsub("[.]", "_", tolower(colnames(iris)))
DBI::dbWriteTable(con, "iris", iris)
var <- "sepal_width"
tbl <- "iris"
num <- 2
val <- "setosa"
glue_sql("
  SELECT {`var`}
  FROM {`tbl`}
  WHERE {`tbl`}.sepal_length > {num}
    AND {`tbl`}.species = {val}
  ", .con = con)

# `glue_sql()` can be used in conjuction with parameterized queries using
# `DBI::dbBind()` to provide protection for SQL Injection attacks
 sql <- glue_sql("
    SELECT {`var`}
    FROM {`tbl`}
    WHERE {`tbl`}.sepal_length > ?
  ", .con = con)
query <- DBI::dbSendQuery(con, sql)
DBI::dbBind(query, list(num))
DBI::dbFetch(query, n = 4)
DBI::dbClearResult(query)

# `glue_sql()` can be used to build up more complex queries with
# interchangeable sub queries. It returns `DBI::SQL()` objects which are
# properly protected from quoting.
sub_query <- glue_sql("
  SELECT *
  FROM {`tbl`}
  ", .con = con)

glue_sql("
  SELECT s.{`var`}
  FROM ({sub_query}) AS s
  ", .con = con)

# If you want to input multiple values for use in SQL IN statements put `*`
# at the end of the value and the values will be collapsed and quoted appropriately.
glue_sql("SELECT * FROM {`tbl`} WHERE sepal_length IN ({vals*})",
  vals = 1, .con = con)

glue_sql("SELECT * FROM {`tbl`} WHERE sepal_length IN ({vals*})",
  vals = 1:5, .con = con)

glue_sql("SELECT * FROM {`tbl`} WHERE species IN ({vals*})",
  vals = "setosa", .con = con)

glue_sql("SELECT * FROM {`tbl`} WHERE species IN ({vals*})",
  vals = c("setosa", "versicolor"), .con = con)
```

# Other implementations

Some other implementations of string interpolation in R (although not using identical syntax).

- [stringr::str_interp](http://stringr.tidyverse.org/reference/str_interp.html)
- [pystr::pystr_format](http://cran.r-project.org/package=pystr)
- [R.utils::gstring](http://cran.r-project.org/package=R.utils)
- [rprintf](http://cran.r-project.org/package=rprintf)
