---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
library(fstrings)
```

# fstrings
[![Travis-CI Build Status](https://travis-ci.org/jimhester/fstrings.svg?branch=master)](https://travis-ci.org/jimhester/fstrings)
[![Coverage Status](https://img.shields.io/codecov/c/github/jimhester/fstrings/master.svg)](https://codecov.io/github/jimhester/fstrings?branch=master)

Python style [f-strings](https://www.python.org/dev/peps/pep-0498/) for R. Small, fast, dependency free interpreted string literals.

## Installation

```{r gh-installation, eval = FALSE}
# install.packages("devtools")
devtools::install_github("jimhester/fstrings")
```

## Example

Long strings can be broken by line and will be concatenated together
```{r}
name <- "Fred"
age <- 50
anniversary <- as.Date("1991-10-12")
f('My name is {name},',
  ' my age next year is {age + 1},',
  ' my anniversary is {format(anniversary, "%A, %B %d, %Y")}.')
```

If you do not want to interpret braces simply double them
```{r}
name <- "Fred"
f("My name is {name}, not {{name}}.")
```

All valid R code works in expressions, including braces and escaping.
Backslashes do need to be doubled just like in all R strings.
```{r}
  `foo}\`` <- "foo"
f("{
      {
        '}\\'' # { and } in comments, single quotes
        \"}\\\"\" # or double quotes are ignored
        `foo}\\`` # as are { in backticks
      }
  }")
```

# Other implementations

Some other implementations of string interpolation in R (although not using the f-string syntax).

- [stringr::str_interp](http://stringr.tidyverse.org/reference/str_interp.html)
- [pystr::pystr_format](https://cran.r-project.org/package=pystr)
- [R.utils::gstring](https://cran.r-project.org/package=R.utils)
