---
title: "Getting started with glue"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with glue}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: 72
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

The glue package contains functions for string interpolation: glueing
together character strings and R code.

To begin, we'll need to load the package.

```{r}
library(glue)
```

## Gluing and interpolating

`glue()` can be used to glue together character strings, such as text:

```{r}
glue("glueing ", "some ", "text ", "together")
```

But it's real power comes with `{}`: anything inside of `{}` will be
evaluated and pasted into the string. This makes it easy to interpolate
variables:

```{r}
name <- "glue"
glue('We are learning how to use the {name} R package.')
```

And more complex functions:

```{r}
release_date <- as.Date("2017-06-13")
glue('The first version of the glue package was released on a {format(release_date, "%A")}.')
```

All valid R code works in expressions, including braces and
escaping.Backslashes do need to be doubled just like in all R strings.
Consider, for instance, adding a variable named `` foo}` `` and printing
its value:

```{r}
`foo}\`` <- "My value"
glue("{
  `foo}\\``
     }")

```

```{r}
`foo}\`` <- "My value"
glue("{
      {
        '}\\'' # { and } in comments, single quotes
        \"}\\\"\" # or double quotes are ignored
        `foo}\\`` # as are { in backticks
      }
  }")
```

## Long lines of text

Glue comes with a number of features that make it easier to use when
work with large quantities of text. Leading whitespace and blank lines
from the first and last lines are automatically trimmed with `glue()`,
letting you indent the strings naturally in code:

```{r}
my_fun <- function() {
  glue("
    A formatted string
    Can have multiple lines
      with additional indention preserved
    ")  
}
my_fun()
```

An add extra newlines can be used if you want a leading or trailing
newline:

```{r}
glue("

  leading or trailing newlines can be added explicitly
  as long as the trailing line contains at least one whitespace character
   
  ")
```

You can use `\\` at the end of a line continues to continue a single
line:

```{r}
glue("
    A formatted string \\
    can also be on a \\
    single line
    ")
```


## Glue in pipes and pipelines

`glue_data()` is useful with magrittr pipes.

```{r}
`%>%` <- magrittr::`%>%`
head(mtcars) %>% glue_data("{rownames(.)} has {hp} hp")
```
`glue()` is useful within dplyr pipelines.

```{r,message=FALSE}
library(dplyr)
head(iris) %>%
  mutate(description = glue("This {Species} has a petal length of {Petal.Length}"))
```


## SQL

glue_sql()\` makes constructing SQL statements safe and easy Use
backticks to quote identifiers, normal strings and numbers are quoted
appropriately for your backend.

```{r}
con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
colnames(iris) <- gsub("[.]", "_", tolower(colnames(iris)))
DBI::dbWriteTable(con, "iris", iris)
var <- "sepal_width"
tbl <- "iris"
num <- 2
val <- "setosa"
glue_sql("
  SELECT {`var`}
  FROM {`tbl`}
  WHERE {`tbl`}.sepal_length > {num}
    AND {`tbl`}.species = {val}
  ", .con = con)
```

`glue_sql()` can be used in conjunction with parameterized queries using
`DBI::dbBind()` to provide protection for SQL Injection attacks

```{r}
sql <- glue_sql("
  SELECT {`var`}
  FROM {`tbl`}
  WHERE {`tbl`}.sepal_length > ?
", .con = con)
query <- DBI::dbSendQuery(con, sql)
DBI::dbBind(query, list(num))
DBI::dbFetch(query, n = 4)
DBI::dbClearResult(query)
```

`glue_sql()` can be used to build up more complex queries with
interchangeable sub queries. It returns `DBI::SQL()` objects which are
properly protected from quoting.

```{r}
sub_query <- glue_sql("
  SELECT *
  FROM {`tbl`}
  ", .con = con)

glue_sql("
  SELECT s.{`var`}
  FROM ({sub_query}) AS s
  ", .con = con)
```

If you want to input multiple values for use in SQL IN statements put
`*` at the end of the value and the values will be collapsed and quoted
appropriately.

```{r}
glue_sql("SELECT * FROM {`tbl`} WHERE sepal_length IN ({vals*})",
  vals = 1, .con = con)

glue_sql("SELECT * FROM {`tbl`} WHERE sepal_length IN ({vals*})",
  vals = 1:5, .con = con)

glue_sql("SELECT * FROM {`tbl`} WHERE species IN ({vals*})",
  vals = "setosa", .con = con)

glue_sql("SELECT * FROM {`tbl`} WHERE species IN ({vals*})",
  vals = c("setosa", "versicolor"), .con = con)
```
