---
title: "Getting started with glue"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Get started with glue}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment = "#>",
                      messages = FALSE,
                      warning = FALSE)

library(glue)
```

The `glue` package contains functions for glueing together character strings and R code.

```{r eval = FALSE}
# Install the released version of the glue package from CRAN:
install.packages("glue")

# Load the glue package
library(glue)
```


## Glueing character strings

`glue()` can be used to glue together character strings, such as text: 
```{r}
glue("glueing", "some", "text", "together")
```

The default for `glue()` is glueing with _no_ separator&mdash; you can provide a separator with the `.sep` argument:
```{r}
glue("glueing", "some", "text", "together", .sep = " ")
glue("glueing", "some", "text", "together", .sep = ".")
```

You can optionally combine strings with `+`:
```{r}
glue("glueing") + " some" + " text" + " together"
```


## Glueing & formatting

Leading whitespace and blank lines from the first and last lines are automatically trimmed with `glue()`, letting you indent the strings naturally in code:
```{r}
glue("
    A formatted string
    Can have multiple lines
      with additional indention preserved
    ")
```

An additional newline can be used if you want a leading or trailing newline.
```{r}
glue("

  leading or trailing newlines can be added explicitly

  ")
```

`\\` at the end of a line continues it without a new line:
```{r}
glue("
    A formatted string \\
    can also be on a \\
    single line
    ")
```


You can add colour to your strings with `glue_col()`:
```{r}
library(crayon)

glue_col("{blue glueing blue text together}")
glue_col("This ", "{red text}", " is not", " {blue blue}")

```


## Glueing character strings _with_ R code

Glue evaluates R code provided in curly braces and glues the result together with character strings.

Variables can be passed directly into strings:
```{r}
name <- "glue"
glue('We are learning how to use the {name} R package.')
```

R expressions can be passed into strings:
```{r}
release_date <- as.Date("2017-06-13")
glue('The first version of the glue package was released on a {format(release_date, "%A")}.')
```


## Glue with a data frame

`glue_data()` can be used to work with a list or data frame:
```{r message=FALSE}
library(dplyr)

glue_data(head(starwars), "{name} is from the planet {homeworld}.")
```

## Glueing with `dplyr` & `ggplot2`

`glue()` can be used with `dplyr` & `ggplot2` pipelines:
```{r, message = FALSE,fig.width=7}
library(dplyr)
library(ggplot2)

starwars %>%
  mutate(new_x_label = glue("{name} ({species})")) %>% 
  filter(height > 200) %>% 
  ggplot(aes(new_x_label, height)) +
  geom_col() +
  coord_flip()
```



## Additional Vignettes

 - [Speed of glue](https://glue.tidyverse.org/articles/speed.html) 
 - [Transformers](https://glue.tidyverse.org/articles/transformers.html)
