---
title: "Transformers"
author: "Jim Hester"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Transformers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Transformers allow you to apply functions to the glue input and output, before
and after evaluation. This allows you to write things like `glue_sql()`, which
automatically quotes variables for you or add a syntax for automatically
collapsing outputs. One important consideration when writing glue wrapper
functions with transformers is to include `.envir = parent.frame()` in the
wrapper to ensure the evaluation environment is correct.

```{r, include = FALSE}
library(glue)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

So you can define a transformer which automatically collapses any glue block ending with `*`.

```{r}
collapse_transformer <- function(regex = "[*]$", ...) {
  function(code, envir) {
    if (grepl(regex, code)) {
        code <- sub(regex, "", code)
    }
    res <- evaluate(code, envir)
    collapse(res, ...)
  }
}

glue("{1:5*}\n{letters[1:5]*}", .transformer = collapse_transformer(sep = ", "))

glue("{1:5*}\n{letters[1:5]*}", .transformer = collapse_transformer(sep = ", ", last = " and "))
```

Or a transformer which converts the text to the equivalent emoji.

```{r, eval = require("emo")}
emoji_transformer <- function(code, envir) {
  if (grepl("[*]$", code)) {
    code <- sub("[*]$", "", code)
    collapse(ji_find(code)$emoji)
  } else {
    ji(code)
  }
}

glue_ji <- function(..., .envir = parent.frame()) {
  glue(..., .open = ":", .close = ":", .envir = .envir, .transformer = emoji_transformer)
}
glue_ji("one :heart:")
glue_ji("many :heart*:")
```

Or a transformer which allows succinct sprintf format strings.
```{r}
sprintf_transformer <- function(code, envir) {
  m <- regexpr(":.+$", code)
  if (m != -1) {
    format <- substring(regmatches(code, m), 2)
    regmatches(code, m) <- ""
    res <- evaluate(code, envir)
    do.call(sprintf, list(glue("%{format}f"), res))
  } else {
    evaluate(code, envir)
  }
}

glue_fmt <- function(..., .envir = parent.frame()) {
  glue(..., .transformer = sprintf_transformer, .envir = .envir)
}
glue_fmt("π = {pi:.2}")
```
