---
title: "Transformers"
author: "Jim Hester"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Transformers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

Transformers allow you to apply functions to the glue input and output, before
and after evaluation. This allows you to write things like `sql_glue()`, which
automatically quotes variables for you or add a syntax for automatically
collapsing outputs.

```{r setup, echo = FALSE}
library(glue)
```

```{r}
sql_quote_transformer <- function(connection) {
  function(code, envir, data) {
    DBI::dbQuoteString(con, evaluate(code, envir, data))
  }
}

con <- DBI::dbConnect(RSQLite::SQLite(), ":memory:")
glue_sql <- function(..., .con) {
  glue(..., .transformer = sql_quote_transformer(.con))
}

var <- "foo"
tbl <- "bar"
glue_sql("SELECT {var} FROM {tbl}", .con = con)
```

Or a transformer which automatically collapses any glue block ending with `*`.

```{r}
collapse_transformer <- function(regex = "[*]$", ...) {
  function(code, envir, data) {
    if (grepl(regex, code)) {
        code <- sub(regex, "", code)
    }
    res <- evaluate(code, envir, data)
    collapse(res, ...)
  }
}

glue("{1:5*}\n{letters[1:5]*}", .transformer = collapse_transformer(sep = ", "))

glue("{1:5*}\n{letters[1:5]*}", .transformer = collapse_transformer(sep = ", ", last = " and "))
```

Or a transformer which converts the text to the equivalent emoji.

```{r, eval = requireNamespace("emo")}
emoji_transformer <- function(code, envir, data) {
  if (grepl("[*]$", code)) {
    code <- sub("[*]$", "", code)
    collapse(emo::ji_find(code)$emoji)
  } else {
    emo::ji(code)
  }
}

glue_ji <- function(...) {
  glue(..., .open = ":", .close = ":", .transformer = emoji_transformer)
}
glue_ji("one :heart:")
glue_ji("many :heart*:")
```

Or a transformer which allows succinct sprintf format strings.
```{r}
sprintf_transformer <- function(code, envir, data) {
  m <- regexpr(":.+$", code)
  if (m != -1) {
    format <- substring(regmatches(code, m), 2)
    regmatches(code, m) <- ""
    res <- evaluate(code, envir, data)
    do.call(sprintf, list(glue("%{format}f"), res))
  } else {
    evaluate(code, envir, data)
  }
}

glue_fmt <- function(...) {
  glue(..., .transformer = sprintf_transformer)
}
glue_fmt("π = {pi:.2}")
```

# TODO: readr::as_chunk_callback
