---
title: "Speed of glue"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
# used inside the plot function, but not in user-visible code
library(forcats)
library(tidyr)
library(ggbeeswarm)
```

```{r setup}
library(glue)
library(ggplot2)
library(bench)
library(dplyr)
```

Glue is advertised as

> Fast, dependency free string literals

So what do we mean when we say that glue is fast?
This does not mean glue is the fastest thing to use in all cases, however for the features it provides we can confidently say it is fast.

A good way to determine this is to compare its speed of execution to some alternatives.

- `base::paste0()`, `base::sprintf()`: Functions in base R implemented in C
  that provide variable insertion (but not interpolation).
- `R.utils::gstring()`: Provides a similar interface as glue, but uses `${}` to
  delimit blocks to interpolate.
- `pystr::pystr_format()`[^1], `rprintf::rprintf()`: Provide an interface
  similar to python string formatters with variable replacement, but not
  arbitrary interpolation.
  
Note: `stringr::str_interp()` was previously included in this benchmark, but is now formally marked as "superseded", in favor of `stringr::str_glue()`, which just calls `glue::glue()`.

```{r setup2, include = FALSE}
plot_comparison <- function(x, ...) {
  dat <- x %>%
    mutate(expression = fct_reorder(as.character(expression), min, .desc = TRUE)) %>%
    arrange(expression) %>%
    mutate(accent = expression == "glue", .after = expression) %>%
    unnest(c(time, gc))
  ggplot(dat) +
    aes(expression, time) +
    geom_quasirandom(aes(color = accent), ..., show.legend = FALSE) +
    scale_color_manual(values = c("FALSE" = "grey", "TRUE" = "orange")) +
    coord_flip()
}
```

## Simple concatenation

```{r, message = FALSE}
bar <- "baz"

simple <- bench::mark(
  glue       = as.character(glue::glue("foo{bar}")),
  gstring    = R.utils::gstring("foo${bar}"),
  paste0     = paste0("foo", bar),
  sprintf    = sprintf("foo%s", bar),
  rprintf    = rprintf::rprintf("foo$bar", bar = bar)
)

simple %>%
  select(expression:total_time) %>%
  arrange(median)

# plotting function defined in a hidden chunk
plot_comparison(simple)
```

While `glue()` is slower than `paste0` and `sprintf()`, it is twice as fast as `gstring()`, and `rprintf()`.

Although `paste0()` and `sprintf()` don't do string interpolation and will likely always be significantly faster than glue, glue was never meant to be a direct replacement for them.

`rprintf::rprintf()` does only variable interpolation, not arbitrary expressions, which was one of the explicit goals of writing glue.

So glue is ~2x as fast as the function (`gstring()`), which has roughly equivalent functionality.

It also is still quite fast, with over 8000 evaluations per second on this machine.

## Vectorized performance

Taking advantage of glue's vectorization is the best way to improve performance.
In a vectorized form of the previous benchmark, glue's performance is much closer to that of `paste0()` and `sprintf()`.

```{r, message = FALSE}
bar <- rep("bar", 1e5)

vectorized <- bench::mark(
  glue    = as.character(glue::glue("foo{bar}")),
  gstring = R.utils::gstring("foo${bar}"),
  paste0  = paste0("foo", bar),
  sprintf = sprintf("foo%s", bar),
  rprintf = rprintf::rprintf("foo$bar", bar = bar)
)

vectorized %>%
  select(expression:total_time) %>%
  arrange(median)

# plotting function defined in a hidden chunk
plot_comparison(vectorized)
```

[^1]: pystr is no longer available from CRAN due to failure to correct
installation errors and was therefore removed from further testing.
